---
title: "Potential explanations for rapid rebound of schistosomiasis following preventive chemotherapy and implications for ongoing control efforts"
author: "Chris Hoover"
bibliography: "references.bib"
csl: "plos.csl"
output:
  word_document:
    toc: true
    toc_depth: 3
geometry: margin = 0.5in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(tidyverse)
require(viridis)
require(sensitivity)
require(parallel)
require(knitcitations)
  cleanbib()
  cite_options(cite.style = "numeric",
               citation_format = "pandoc")
devtools::load_all()
```

# Background

## Highly-dispersed, age-dependent infection intensities
Infection intensity of schistosomes and other helminths is highly variable and generally well-characterized by a negative binomial distribution in which the clumping parameter, $\kappa$, is well below 1, implying a highly overdispersed distribution in which few individuals host many parasites. Furthermore, infection intensities tend to be age-dependent, with intensity peaking among school-aged children (SAC), and declining in older individuals. A number of phenomenon could produce these observed patterns of infection.

## Rapid rebound of infection following preventive chemotherapy/mass drug administration in some communities
While many communities receiving preventive chemotherapy exhibit the desired result of reductions in transmission and infection intensity others rebound to pre-treatment levels of infection prior to the next round of PC, often within a year. Still others even experience increases in transmission intensity or prevalence.

# Potential explanations for rapid rebound and overdispersion

## Infection intensity is determined by exposure


## Acquired immunity regulates parasite burdens


## Parasite crowding restricts egg output


## High snail-to-human transmission
Conventional wisdom that rebound is proportional to inverse of adult parasite lifespan, but this ignores critical dynamics of snail-to-human transmission. If infected snail density is high enough and persists long enough after human treatment to reinfect individuals, rapid rebound is likely.


```{r}
data.frame("Model 1" = c("X","","","","100"),
           "Model 2" = c("","X","","","100"),
           "Model 3" = c("","","X","","100"),
           "Model 4" = c("","","","X","100"),
           "Model 5" = c("X","X","","","100"),
           "Model 6" = c("X","","X","","100"),
           "Model 7" = c("X","","","X","100"),
           "Model 8" = c("","X","X","","100"),
           "Model 9" = c("","X","","X","100"),
           "Model 10" = c("","","X","X","100"),
           "Model 11" = c("X","X","X","","100"),
           "Model 12" = c("X","X","","X","100"),
           "Model 13" = c("X","","X","X","100"),
           "Model 14" = c("","X","X","X","100"),
           "Model 15" = c("X","X","X","X","100")) %>% t() %>%
  knitr::kable(col.names = c("Exposure", "Immunity", "Crowding", "Snail-to-man", "Model Fit"))
```

# Figures
## Fit of different models to "responder" and "persistent hotspot" communities
