---
title: "Garchitorena et al model optimize across {$R_0$, P} parameter space"
output: 
  github_document:
    toc: true
    toc_depth: 3
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(viridis)
require(furrr)
require(tidyverse)

devtools::load_all("../../DDNTD")

#Define function for equations in github_documents from https://github.com/STAT545-UBC/Discussion/issues/102
latexImg = function(latex){

    link = paste0('http://latex.codecogs.com/gif.latex?',
           gsub('\\=','%3D',URLencode(latex)))

    link = gsub("(%..)","\\U\\1",link,perl=TRUE)
    return(paste0('![](',link,')'))
}
```

# Garchitorena et al simplified model  
Using a simple, generalizable model of NTD transmission presented in [Garchitorena et al](http://rstb.royalsocietypublishing.org/content/372/1722/20160128):

The model consists of two state variables, $I$ and $W$, that correspond to the prevalence of infection in the human population (i.e. proportion infected at time=$t$) and the degree of contamination of the environment with the disease-causing agent, respectively. We make the simplifying assumption that individuals can only be susceptible, $S$, or infected, $I$, meaning $S+I=1$ and eliminating the need for a recovered $R$ compartment as is typical of SIR models but would complicate things here. The model equations then are:  

`r latexImg("\\frac{dI}{dt}=(\\beta_EW+\\beta_DI)(1-I)-\\gamma I") `

`r latexImg("\\frac{dW}{dt}=\\Omega+V\\sigma\\lambda I-\\rho W") `

For most environmentally mediated infectious diseases, transmission to humans is exclusively from the environment and there is no exogenous production of infectious agents in the environment (i.e. $\Omega=0$ and $\beta_D=0$. We can derive $R_0$ for this system quite simply since transmission is only determined by the environmental component: 

`r latexImg("R_0=\\frac{V\\sigma\\lambda\\beta_E}{\\gamma\\delta}")`

with parameter definitions and values in table 1, we have $R_0\approx$ `r round(garch_r0(garch_pars), 1)`

```{r table1, echo = FALSE}
tab1 <- data.frame(val = as.character(round(garch_pars,5)), 
                   des = c("Transmission rate from environment to human population",
                           "Human to human transmission rate",
                           "Rate of recovery from infected back to susceptible",
                           "Recruitment rate of infectious agents in the environment",
                           "Abundance of vectors/intermediate hosts/suitable environment",
                           "Recruitment rate of infectious agents by infectious individuals",
                           "Fraction of infectious agents produced that reach the environment",
                           "Mortality rate of infectious agents in the environment"))

rownames(tab1) <- c("$\\beta_E$","$\\beta_D$","$\\gamma$","$\\Omega$",
                     "$V$","$\\lambda$","$\\sigma$","$\\rho$")

knitr::kable(tab1, row.names = TRUE, 
             col.names = c("Value", "Description"), 
             format = "markdown", escape = FALSE, 
             caption = "Parameter values and descriptions used in the model")
```

## Example Dynamics  
### Without dd  
```{r mod_example}
#Get equilibrium estimates for state variables
garch_eq <- runsteady(y = c(I = 0.5, W = 20), func = DDNTD::garch_mod,
                      parms = DDNTD::garch_pars)[["y"]]

#Events representing interventions to implement in the model
n_years <- 20                  #Simulate 10 years of annual intervention
drug_efficacy <- 0.75            #75% reduction in prevalence at each drug treatment
run_time <- c(1:(365*n_years)) #Run model for 10 years

#data frame with event times for drug administration
drugs <- data.frame(var=rep('I', times = n_years/2),
                    time = c(1:(n_years/2))*365,
                    value = rep((1-drug_efficacy), times = n_years/2),
                    method = rep("mult", times = (n_years/2)))
    
#Run model with annual drug administration intervention
drug_only <- sim_garch_mod(garch_eq, run_time, garch_mod, garch_pars,
                           events_df = drugs)

#Reduce vector/intermediate host population
pars_env <- garch_pars
pars_env["V"] <- garch_pars["V"] * 0.8 

#Run model with annual drug administration and environmental intervention 
drug_env <- sim_garch_mod(garch_eq, run_time, garch_mod, pars_env,
                           events_df = drugs)

#Plot results
rbind(drug_env, drug_only) %>% 
  mutate(Intervention = c(rep("Drug + Env", length(run_time)),
                          rep("Drug", length(run_time)))) %>% 
  ggplot(aes(x = time/365, y = I, lty = Intervention)) + 
    theme_classic() + theme(legend.position = c(0.5,0.85)) +
    geom_line(size = 0.75) + 
  labs(x = "time (years)", 
       y = "Prevalence",
       title = "Model dynamics with annual MDA",
       subtitle = "with and without environmental intervention")
  
```

### With dd  
```{r dd_mod_reff}
#Get equilibrium estimates for state variables
garch_eq_dd <- runsteady(y = c(I = 0.5, W = 20), func = DDNTD::garch_mod_dd,
                      parms = DDNTD::garch_pars)[["y"]]

plot(seq(0,1,0.01), sapply(seq(0,1,0.01), function(I) garch_r0(garch_pars) * (4*I*(1-I))),
     xlab = "Prevalence (I)", ylab = expression(R[eff]), type = "l")

```

```{r example_dd}
#Run model with annual drug administration intervention
drug_only_dd <- sim_garch_mod(garch_eq, run_time, garch_mod_dd, garch_pars,
                           events_df = drugs)

#Run model with annual drug administration and environmental intervention 
drug_env_dd <- sim_garch_mod(garch_eq, run_time, garch_mod_dd, pars_env,
                           events_df = drugs)

#Plot results
rbind(drug_env, drug_only) %>% 
  mutate(Intervention = c(rep("Drug + Env", length(run_time)),
                          rep("Drug", length(run_time)))) %>% 
  ggplot(aes(x = time/365, y = I, lty = Intervention)) + 
    theme_classic() + theme(legend.position = c(0.5,0.85)) +
    geom_line(size = 0.75) + 
  labs(x = "time (years)", 
       y = "Prevalence",
       title = "Model dynamics with annual MDA",
       subtitle = "with and without environmental intervention")

```

# Optimal control framework  

```{r opt_setup}
#Intervention costs, available capital, and conversion parameters
H <- 1000      # number of people
M <- 1.5*H     # available capital, here modeled as $1.50 per person

theta <- 1.8   #cost of 1% reduction in prevalence
mu <- theta    #cost of 1% reduction in W*
  
  
#Parameters for the model and utility function: 
mdp_pars <- c(garch_pars, 
              "d" = 100,      # Arbitrary cost of having prevalence of I_t
              "M" = M,        # Capital available to spend on MDA
              "theta" = theta, # Scaling of capital spent on MDA to reduction in prevalence, 
              "mu" = mu,    # Scaling of environmental intervention to cost (impact per cost)
              "delta" = 0.95)  #Arbitrary discounting rate  

test_grid <- expand.grid(R0 = c(1:10),
                         script_P = exp(seq(log(0.01), log(100), length.out = 20)),
                         T_frame = c(2, 10, 20)*365,
                         freq = 365,
                         M = c(500, 1000, 1500),
                         A = seq(0,1,0.025))
```

## Identify optimal control strategy with no density dependence  
```{r opt_no_dd, cache = T}
plan(multiprocess)
#tic()
test_Us <- future_pmap_dbl(as.list(test_grid), sim_w_inputs, pars = mdp_pars, mod = garch_mod)
#toc()

test_fin <- cbind(test_grid, test_Us)

```

```{r opt_no_dd_plot}
test_fin %>% 
  group_by(R0, script_P, T_frame, M) %>% 
  summarise(U = max(test_Us),
            A_opt = A[which(test_Us == U)]) %>% 
  mutate(C_labs = paste0("C = ", M),
         C_labs = factor(C_labs, levels = c("C = 500", "C = 1000", "C = 1500")),
         T_labs = paste0("T = ", T_frame/365),
         T_labs = factor(T_labs, levels = c("T = 2", "T = 10", "T = 20"))) %>% 
  ggplot(aes(x = R0, y = script_P, z = A_opt)) +
    facet_grid(C_labs ~ T_labs) +
    geom_tile(aes(fill = A_opt)) +
    theme_classic() +
    scale_fill_viridis(option = "magma", direction = 1) +
    scale_x_continuous(breaks = c(1, 5, 10)) +
    scale_y_continuous(trans = "log",
                       breaks = c(0.01,0.1,1,10,100),
                       labels = c(0.01,0.1,1,10,100)) +
    labs(x = expression(R[0]),
         y = expression(italic(Rho)),
         fill = expression(italic(A)))

```


## Identify optimal control strategy with density dependence  

```{r opt_dd, cache = T}
plan(multiprocess)
#tic()
test_Us_dd <- future_pmap_dbl(as.list(test_grid), sim_w_inputs, pars = mdp_pars, mod = garch_mod_dd)
#toc()

test_fin_dd <- cbind(test_grid, test_Us_dd)
```

```{r opt_dd_plot}
test_fin_dd %>% 
  group_by(R0, script_P, T_frame, M) %>% 
  summarise(U = max(test_Us_dd),
            A_opt = A[which(test_Us_dd == U)]) %>% 
  ggplot(aes(x = R0, y = script_P, z = A_opt)) +
    facet_grid(M ~ T_frame) +
    geom_tile(aes(fill = A_opt)) +
    theme_classic() +
    scale_fill_viridis(option = "magma", direction = 1) +
    scale_y_continuous(trans = "log",
                       breaks = c(0.01,0.1,1,10,100),
                       labels = c(0.01,0.1,1,10,100)) +
    labs(x = expression(R[0]),
         y = expression(italic(Rho)),
         fill = expression(italic(A)))

```

